Improving real-time trend estimates using local parametrisation of local polynomial regression filters

### 2024-02-12 article_jos.qmd
numbered section
R logo removed

# Introduction {-}

@@

Using local polynomial regression, @proietti2008 developed general class of symmetric and asymmetric moving averages.
In particular, it makes possible to reproduce Henderson's symmetric filter and Musgrave's asymmetric filters.
However, these methods have two drawbacks.
Firstly, the phase shift (i.e. the delay in detecting turning points) is not directly controlled.
Secondly, asymmetric filters depend on a parameter that is generally estimated over the whole series using assumptions that may be wrong locally.
The aim of this study is to propose extensions to the @proietti2008 class of filters in order to take these shortcomings into account.
The extensions have been implemented in the {{< fa brands r-project >}} `rjd3filters` package (available at <https://github.com/rjdemetra/rjd3filters>) and are therefore easily reusable.

@@

@ch15HBSA propose a general approach to derive linear filters, based on an optimisation problem involving three criteria: *Fidelity* ($F_g$), *Smoothness* ($S_g$) and *Timeliness* ($T_g$).
Fidelity can be directly linked to the reduction in variance created by the filter, and timeliness to the notion of phase shift, which again we want to be low. 
This method can be used in{{< fa brands r-project >}} with the function `rjd3filters::fst_filter()`.

@trilemmaWMR2019 propose an approach based on the decomposition of the mean squared error between the symmetric and the asymmetric filter in four quantities: *Accuracy* ($A_w$), *Timeliness* ($T_w$), *Smoothness* ($S_w$) and *Residual* ($R_w$).
A simplified version of this method can be used in {{< fa brands r-project >}} with the function `rjd3filters::dfa_filter()`.

@@

In practice, this means assuming that the trend is locally polynomial. 
Various estimation methods can be used to derive symmetrical and asymmetrical moving averages.  
@GrayThomson1996 propose a complete statistical framework which makes it possible, in particular, to model the error in approximating the trend by local polynomials.
However, as the specification of this error is generally complex, simpler models may be preferred, such as that of @proietti2008.  
Finally, @dagumbianconcini2008 propose a similar modelling of the trend-cycle but using the theory of Hilbert spaces with reproducing kernels for estimation, which has the particular advantage of facilitating the calculation of different moving averages at different time frequencies.
This method can be used in{{< fa brands r-project >}} with the function `rjd3filters::rkhs_filter()`.

@@

Using the same notation as in @sec-lppasymf, $\boldsymbol\theta$ represents the symmetric filter and $\boldsymbol v$ the asymmetric filter.
With $\boldsymbol\theta=\begin{pmatrix}\boldsymbol\theta_p\\\boldsymbol\theta_f\end{pmatrix}$ and $\boldsymbol\theta_p$ of the same length of $\boldsymbol v$, and $\boldsymbol g=\boldsymbol v-\boldsymbol \theta_p$, the timeliness criterion can be expressed as:
$$
T_g(\boldsymbol v)=\boldsymbol v'\boldsymbol T\boldsymbol v=\boldsymbol g'\boldsymbol T\boldsymbol g+2\boldsymbol \theta_p'\boldsymbol T\boldsymbol g+\boldsymbol \theta_p'\boldsymbol T\boldsymbol \theta_p
\quad\text{ with }\boldsymbol T\text{ a symmetric matrix}.
$$
Furthermore, the objective function $\varphi$ of @eq-lppasym can be rewritten:
\begin{align*}
\varphi(\boldsymbol v)&=(\boldsymbol v-\boldsymbol \theta_p)'\boldsymbol D_{p}(\boldsymbol v-\boldsymbol \theta_p)+
  \boldsymbol \theta_f'\boldsymbol D_{f}\boldsymbol \theta_f+
  [\boldsymbol \delta'(\boldsymbol Z_{p}'\boldsymbol v-\boldsymbol Z'\boldsymbol \theta)]^{2}+
2\boldsymbol l'(\boldsymbol U_{p}'\boldsymbol v-\boldsymbol U'\boldsymbol \theta)\\
&=\boldsymbol g'\boldsymbol Q\boldsymbol g-2\boldsymbol P\boldsymbol g+2\boldsymbol l'(\boldsymbol U_{p}'\boldsymbol v-\boldsymbol U'\boldsymbol \theta)+\boldsymbol c\quad\text{ with }\boldsymbol v
\begin{cases}
\boldsymbol Q=\boldsymbol D_p+\boldsymbol Z_p\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z'_p \\
\boldsymbol P=\boldsymbol \theta_f\boldsymbol Z_f\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z_p'\\
\boldsymbol c\text{ a constant independent of }\boldsymbol v
\end{cases}.
\end{align*}
Adding the timeliness criterion:
$$
\widetilde\varphi(\boldsymbol v)=\boldsymbol g'\widetilde {\boldsymbol Q}\boldsymbol g-
2\widetilde {\boldsymbol P}\boldsymbol g+2\boldsymbol l'(\boldsymbol U_{p}'\boldsymbol v-\boldsymbol U'\boldsymbol \theta)+
\widetilde {\boldsymbol c}\quad\text{with }
\begin{cases}
\widetilde {\boldsymbol Q}=\boldsymbol D_p+\boldsymbol Z_p\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z'_p + \alpha_T\boldsymbol T\\
\widetilde {\boldsymbol P}=\boldsymbol \theta_f\boldsymbol Z_f\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z_p'-\alpha_T\boldsymbol \theta_p\boldsymbol T\\
\widetilde {\boldsymbol c}\text{ a constant independent of }\boldsymbol v
\end{cases},
$$
where $\alpha_T$ is the weight associated to the timeliness criterion. 
With $\alpha_T=0$ we find $\varphi(\boldsymbol v)$.
This extension therefore makes it possible to find all the symmetric and asymmetric filters presented in the previous section but also generalizes the approach of @GrayThomson1996.
This extension is implemented in the {{< fa brands r-project >}} function `rjd3filters::lp_filter()`.

@@

# Conclusion {.unnumbered}

@@

This paper presents the {{< fa brands r-project >}} package `rjd3filters`, which implements a several methods to build moving average for real-time trend-cycle estimation.
It also offers various functions for the theoretical analysis of moving averages (gain functions, phase shift, quality criteria, etc.) and judging the quality of the latest estimates (for example with implicit forecasts).
It also makes it easy to combine moving averages and also to integrate custom moving averages into the X-11 seasonal adjustment algorithm (with the `rjd3filters::x11plus_trend()` function).
`rjd3filters` thus facilitates research into the use of moving averages as part of real-time cycle trend estimation.
This is the case, for example, with the local parametrisation of Musgrave moving averages and other asymmetric moving averages linked to local polynomial regression, presented in this article.

@@

In this appendix we present the {{< fa brands r-project >}} code that can be used to estimate trend-cycle components for the US Employment with local polynomial filters.

>>>>

# Introduction

@@

Using local polynomial regression, @proietti2008 developed general class of symmetric and asymmetric moving averages.
In particular, it makes possible to reproduce Henderson's symmetric filter and Musgrave's asymmetric filters.
However, these methods have two drawbacks.
Firstly, the phase shift (i.e. the delay in detecting turning points) is not directly controlled.
Secondly, asymmetric filters depend on a parameter that is generally estimated over the whole series using assumptions that may be wrong locally.
The aim of this study is to propose extensions to the @proietti2008 class of filters in order to take these shortcomings into account.
The extensions have been implemented in the R `rjd3filters` package (available at <https://github.com/rjdemetra/rjd3filters>) and are therefore easily reusable.

@@

@ch15HBSA propose a general approach to derive linear filters, based on an optimisation problem involving three criteria: *Fidelity* ($F_g$), *Smoothness* ($S_g$) and *Timeliness* ($T_g$).
Fidelity can be directly linked to the reduction in variance created by the filter, and timeliness to the notion of phase shift, which again we want to be low. 
This method can be used inR with the function `rjd3filters::fst_filter()`.

@trilemmaWMR2019 propose an approach based on the decomposition of the mean squared error between the symmetric and the asymmetric filter in four quantities: *Accuracy* ($A_w$), *Timeliness* ($T_w$), *Smoothness* ($S_w$) and *Residual* ($R_w$).
A simplified version of this method can be used in R with the function `rjd3filters::dfa_filter()`.

@@

In practice, this means assuming that the trend is locally polynomial. 
Various estimation methods can be used to derive symmetrical and asymmetrical moving averages.  
@GrayThomson1996 propose a complete statistical framework which makes it possible, in particular, to model the error in approximating the trend by local polynomials.
However, as the specification of this error is generally complex, simpler models may be preferred, such as that of @proietti2008.  
Finally, @dagumbianconcini2008 propose a similar modelling of the trend-cycle but using the theory of Hilbert spaces with reproducing kernels for estimation, which has the particular advantage of facilitating the calculation of different moving averages at different time frequencies.
This method can be used inR with the function `rjd3filters::rkhs_filter()`.

@@

Using the same notation as in @sec-lppasymf, $\boldsymbol\theta$ represents the symmetric filter and $\boldsymbol v$ the asymmetric filter.
With $\boldsymbol\theta=\begin{pmatrix}\boldsymbol\theta_p\\\boldsymbol\theta_f\end{pmatrix}$ and $\boldsymbol\theta_p$ of the same length of $\boldsymbol v$, and $\boldsymbol g=\boldsymbol v-\boldsymbol \theta_p$, the timeliness criterion can be expressed as:
$$
T_g(\boldsymbol v)=\boldsymbol v'\boldsymbol T\boldsymbol v=\boldsymbol g'\boldsymbol T\boldsymbol g+2\boldsymbol \theta_p'\boldsymbol T\boldsymbol g+\boldsymbol \theta_p'\boldsymbol T\boldsymbol \theta_p
\quad\text{ with }\boldsymbol T\text{ a symmetric matrix}.
$$
Furthermore, the objective function $\varphi$ of @eq-lppasym can be rewritten:
\begin{align*}
\varphi(\boldsymbol v)&=(\boldsymbol v-\boldsymbol \theta_p)'\boldsymbol D_{p}(\boldsymbol v-\boldsymbol \theta_p)+
  \boldsymbol \theta_f'\boldsymbol D_{f}\boldsymbol \theta_f+
  [\boldsymbol \delta'(\boldsymbol Z_{p}'\boldsymbol v-\boldsymbol Z'\boldsymbol \theta)]^{2}+
2\boldsymbol l'(\boldsymbol U_{p}'\boldsymbol v-\boldsymbol U'\boldsymbol \theta)\\
&=\boldsymbol g'\boldsymbol Q\boldsymbol g-2\boldsymbol P\boldsymbol g+2\boldsymbol l'(\boldsymbol U_{p}'\boldsymbol v-\boldsymbol U'\boldsymbol \theta)+\boldsymbol c\quad\text{ with }\boldsymbol v
\begin{cases}
\boldsymbol Q=\boldsymbol D_p+\boldsymbol Z_p\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z'_p \\
\boldsymbol P=\boldsymbol \theta_f\boldsymbol Z_f\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z_p'\\
\boldsymbol c\text{ a constant independent of }\boldsymbol v
\end{cases}.
\end{align*}
Adding the timeliness criterion:
$$
\widetilde\varphi(\boldsymbol v)=\boldsymbol g'\widetilde {\boldsymbol Q}\boldsymbol g-
2\widetilde {\boldsymbol P}\boldsymbol g+2\boldsymbol l'(\boldsymbol U_{p}'\boldsymbol v-\boldsymbol U'\boldsymbol \theta)+
\widetilde {\boldsymbol c}\quad\text{with }
\begin{cases}
\widetilde {\boldsymbol Q}=\boldsymbol D_p+\boldsymbol Z_p\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z'_p + \alpha_T\boldsymbol T\\
\widetilde {\boldsymbol P}=\boldsymbol \theta_f\boldsymbol Z_f\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z_p'-\alpha_T\boldsymbol \theta_p\boldsymbol T\\
\widetilde {\boldsymbol c}\text{ a constant independent of }\boldsymbol v
\end{cases},
$$
where $\alpha_T$ is the weight associated to the timeliness criterion. 
With $\alpha_T=0$ we find $\varphi(\boldsymbol v)$.
This extension therefore makes it possible to find all the symmetric and asymmetric filters presented in the previous section but also generalizes the approach of @GrayThomson1996.
This extension is implemented in the R function `rjd3filters::lp_filter()`.

@@

# Conclusion

@@

This paper presents the R package `rjd3filters`, which implements a several methods to build moving average for real-time trend-cycle estimation.
It also offers various functions for the theoretical analysis of moving averages (gain functions, phase shift, quality criteria, etc.) and judging the quality of the latest estimates (for example with implicit forecasts).
It also makes it easy to combine moving averages and also to integrate custom moving averages into the X-11 seasonal adjustment algorithm (with the `rjd3filters::x11plus_trend()` function).
`rjd3filters` thus facilitates research into the use of moving averages as part of real-time cycle trend estimation.
This is the case, for example, with the local parametrisation of Musgrave moving averages and other asymmetric moving averages linked to local polynomial regression, presented in this article.

@@

In this appendix we present the R code that can be used to estimate trend-cycle components for the US Employment with local polynomial filters.
